Accuracy and precision of zero-heat-flux temperature monitoring

### 2020-03-18 manuscript/manuscript.Rmd
table 1 updated

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring"
subtitle: A systematic review and meta-analysis
description: "A systematic review and meta-analysis"
# author:
# - name: Aaron Conway
#   affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
#   email: aaron.conway@utoronto.ca
#   url: https://www.aaronconway.info
# - name: Megan Bittner
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
# - name: Dan Phan
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
# - name: Navpreet Kamboj
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
#   email: navpreet.kamboj@mail.utoronto.ca
# - name: Kristina Chang
#   affiliation: Toronto General Hospital, UHN
# - name: Elizabeth Tipton
#   affiliation: Department of Statistics, Northwestern University
# - name: Matteo Parotto
#   affiliation: Toronto General Hospital, UHN
# date: "`r Sys.Date()`"
# repository_url: https://github.com/awconway/zhf-review
# twitter:
#   creator: "@aw_conway"
# output: distill::distill_article
bibliography: Extras/bibliography.bib
csl: Extras/anesthesiology.csl
author:
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Megan Bittner BSc:
      institute: [LBFON]
      correspondence: false
  - Dan Phan BSc:
      institute: [LBFON]
      correspondence: false
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN:
      institute: [LBFON]
      correspondence: false
  - Elizabeth Tipton PhD:
      institute: [NW]
      correspondence: false
  - Matteo Parotto MD, PhD:
      institute: [TGH, UoT]
      correspondence: false
institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - NW: Department of Statistics, Northwestern University, Evanston, IL, United States
  - TGH: Department of Anesthesia and Pain Management, Toronto General Hospital, UHN, Toronto, Canada
  - UoT: Department of Anesthesia and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
output:
  bookdown::word_document2:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
always_allow_html: true
params:
  distill: FALSE
  word: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=FALSE,
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggprisma) # https://github.com/awconway/ggprisma
library(kableExtra)
library(knitr)
library(magrittr)
source("R/loa_maker.R")
source("R/meta.R")
```
\pagebreak

@@

@@

```{r meta-analysis-prep, include=FALSE}
data <- data %>%
  dplyr::mutate(c = N/n)

@@

@@

Medline and EMBASE were searched for studies that reported on a measurement of core or peripheral temperature that coincided with a measurement from the zero-heat-flux device. Study selection and quality assessment (using QUADAS-2) was performed independently. The GRADE approach was used to summarize the strength of the evidence.  Pooled estimates of the mean bias and limits of agreement with outer 95% confidence intervals (population limits of agreement) were calculated.

@@

Published studies were found by searching Medline and EMBASE to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located at this [link](https://aconway.dev/zhf). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses.[@R2018] All data and R code used in the meta-analyses can be located at this [link](https://github.com/awconway/zhf-review).

@@

```{r fig-1,  fig.cap="PRISMA Flow Diagram", echo=FALSE, fig.width=5.5, fig.path="figures/", include=params$distill, eval=params$distill}
# https://github.com/awconway/ggprisma
ggprisma::ggprisma(retrieved = 130, included = 16, duplicates = 35, full_text = 22, wrong_intervention = 4, wrong_comparator = 2, wrong_design = 1, awaiting_classification = 2)
```

The characteristics of included studies can be viewed at this [link](https://aconway.dev/zhf). The primary comparison of zero-heat-flux versus core temperature measurements (eg, arterial, bladder, esophageal or rectal) consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. In total, data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included in this comparison (two studies did not report the total number of measurements included in their analysis). The sensitivity analysis for the primary comparison with only studies that were judged as low risk of bias across all domains included `r low_risk$Comparisons` comparisons from `r low_risk$Studies` studies that enrolled `r low_risk$Participants` participants with `r low_risk$Measurements` paired measurements (two studies did not report the total number of measurements included in their analysis). There were `r ICU$Studies` studies that compared zero-heat-flux to core temperature measurements in ICU patients, comprising `r ICU$Comparisons` comparisons with `r ICU$Measurements` measurements from `r ICU$Participants` participants. Another `r OT$Studies` studies were included that compared zero-heat-flux to core temperature measurements in patients undergoing surgery, comprising `r OT$Comparisons` comparisons with `r OT$Measurements` measurements from `r OT$Participants` participants. Nasopharyngeal thermometers were used as the comparator devices in `r NPA$Studies` studies, with `r NPA$Measurements` paired measurements from `r NPA$Measurements` participants. Sublingual thermometers were used as the comparator devices in `r SL$Studies` studies that reported `r SL$Measurements` paired measurements from `r SL$Participants` participants.

All studies included in this systematic review evaluated the zero-heat-flux temperature monitoring system manufactured by 3M. Previously known as the SpotOn Temperature Monitoring System, the 3M zero-heat-flux device is now referred to commercially as the Bair Hugger Temperature Monitoring System. All comparisons reported adherence to the zero-heat-flux device manufacturer instructions and placed the sensor on the forehead of the participants. One included study also reported results for comparisons where the zero-heat-flux thermometer was not placed on the forehead. We did not include these comparisons in our meta-analysis.

@@

A graphical overview of the findings for the primary analysis is presented in Figure 2. The pooled estimate for the mean bias between zero-heat-flux and core temperature measurements was `r primary$bias_mean`°C. The population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` studies). The mean bias was again `r low_risk$bias_mean`°C with population limits of agreement spanning from `r low_risk$CI_L_rve`°C to `r low_risk$CI_U_rve`°C. A further sensitivity analysis excluding studies that received industry funding revealed population limits of agreement that were considerably wider than the primary analysis (`r conflict$CI_L_rve`°C to `r conflict$CI_U_rve`°C). The GRADE rating for this sensitivity was downgraded to low quality though, due to concerns about study limitations, inconsistency and imprecision.

```{r fig-2, preview=TRUE, layout="l-screen-inset", fig.align="center", fig.cap="Graphical overview of findings for the primary analysis", echo=FALSE, fig.path="", include=params$distill}
knitr::include_graphics("figures/gofer_core.png")
```

Table 1 presents results of the primary, sensitivity and all subgroup analyses for the primary comparison as well as secondary comparisons. We conducted two subgroup analyses for the primary comparison according to the clinical setting in which the study was conducted. In the subset of studies conducted in the ICU, the mean bias was `r ICU$bias_mean`°C with population limits of agreement between `r ICU$CI_L_rve`°C and `r ICU$CI_U_rve` °C. The GRADE rating for quality of evidence was downgraded to very low quality due to study limitations, inconsistency and imprecision. In the subset of studies that evaluated the use of zero-heat-flux thermometers during surgery found a mean bias of `r OT$bias_mean`°C with population limits of agreement from `r OT$CI_L_rve`°C to `r OT$CI_U_rve`°C. The quality of evidence was rated as moderate quality due to concerns about study limitations.

@@

Zero-heat-flux thermometers were compared with sublingual thermometers and nasopharyngeal thermometers in the studies included in this review. The mean bias between zero-heat-flux and sublingual temperature measurements in meta-analysis of results from `r SL$Studies` studies was `r SL$bias_mean`°C. Due to the limited number of studies and measurements, population limits of agreement were extremely wide, spanning from `r SL$CI_L_rve`°C to `r SL$CI_U_rve`°C. The quality of evidence for this comparison was rated as very low quality due to serious concerns about imprecision.

The mean bias between zero-heat-flux and nasopharyngeal temperature measurements in meta-analysis of results from `r NPA$Studies` studies was `r NPA$bias_mean`°C. Population limits of agreement were `r NPA$CI_L_rve`°C to `r NPA$CI_U_rve`°C. We downgraded the quality of evidence to low, again due to concerns about imprecision and study limitations.

@@

It should be noted that we did not include studies that used the Temple Touch Pro because it is not strictly a zero-heat-flux device. This is a new thermometer that is similar to the zero-heat-flux device in that it is placed cutaneously, but the underlying technology is different. Far fewer studies have  evaluated the Temple Touch Pro compared to the zero-heat-flux device.[@evron2017evaluation; @sastre2019evaluation]

@@

Figure 2: Graphical overview of findings for the primary analysis")
```

>>>>

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring"
subtitle: A systematic review and meta-analysis
# description: "A systematic review and meta-analysis"
# author:
# - name: Aaron Conway
#   affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
#   email: aaron.conway@utoronto.ca
#   url: https://www.aaronconway.info
# - name: Megan Bittner
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
# - name: Dan Phan
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
# - name: Navpreet Kamboj
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
#   email: navpreet.kamboj@mail.utoronto.ca
# - name: Kristina Chang
#   affiliation: Toronto General Hospital, UHN
# - name: Elizabeth Tipton
#   affiliation: Department of Statistics, Northwestern University
# - name: Matteo Parotto
#   affiliation: Toronto General Hospital, UHN
# date: "`r Sys.Date()`"
# repository_url: https://github.com/awconway/zhf-review
# twitter:
#   creator: "@aw_conway"
# output: distill::distill_article
bibliography: Extras/bibliography.bib
csl: Extras/anesthesia-and-analgesia.csl
author:
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Megan Bittner BSc:
      institute: [LBFON]
      correspondence: false
  - Dan Phan BSc:
      institute: [LBFON]
      correspondence: false
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN:
      institute: [LBFON]
      correspondence: false
  - Elizabeth Tipton PhD:
      institute: [NW]
      correspondence: false
  - Matteo Parotto MD, PhD:
      institute: [TGH, UoT]
      correspondence: false
institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - NW: Department of Statistics, Northwestern University, Evanston, IL, United States
  - TGH: Department of Anesthesia and Pain Management, Toronto General Hospital, UHN, Toronto, Canada
  - UoT: Department of Anesthesia and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
output:
  bookdown::word_document2:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
always_allow_html: true
params:
  distill: FALSE
  word: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=FALSE,
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggprisma) # https://github.com/awconway/ggprisma
library(kableExtra)
library(knitr)
library(magrittr)
source("R/loa_maker.R")
source("R/meta.R")
```

@@

Author contributions: 
AC: Study design, analysis, wrote manuscript; MB: Study design, data extraction, risk of bias assessment, data analysis, contributed to writing manuscript; DP: Study design, data extraction, risk of bias assessment, revised the manuscript for important intellectual content; NK: Study design, revised the manuscript for important intellectual content; KC: Study design, revised the manuscript for important intellectual content; ET: Study design, data analysis, revised the manuscript for important intellectual content; MP: Study design, revised the manuscript for important intellectual content.

**Glossary of terms:**

GRADE = Grading of Recommendations, Assessment, Development and Evaluations

QUADAS-2 = Revised Quality Assessment of Diagnostic Accuracy Studies tool

LoA = Limits of Agreement

CI = Confidence intervals

SD^2^ = Variance

&tau;^2^ = Tau-squared

@@

```{r meta-analysis-prep, include=FALSE}

data <- data %>%
  dplyr::mutate(c = N/n)

@@

data_PA <- data_core %>% 
   dplyr::filter(comparison=="PA")

@@

Medline and EMBASE were searched for studies that reported on a measurement of core or peripheral temperature that coincided with a measurement from the zero-heat-flux device. Study selection and quality assessment was performed independently using the Revised Quality Assessment of Diagnostic Accuracy Studies tool (QUADAS-2). The Grading of Recommendations, Assessment, Development and Evaluations (GRADE) approach was used to summarize the strength of the evidence.  Pooled estimates of the mean bias and limits of agreement with outer 95% confidence intervals (population limits of agreement) were calculated.

@@

Published studies were found by searching Medline and EMBASE to January 2020. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers.[@eshraghi2014] The search strategies used for each database are available on a [webpage](https://aconway.dev/zhf) containing supplementary information for the systematic review. Selection of studies was undertaken independently by two reviewers.

@@

Measures of uncertainty were included in our meta-analyses by calculating the outer 95% confidence intervals for pooled limits of agreement. We also accounted for repeated measurements if they were not properly adjusted for in individual studies. This was achieved by using weights proportional to the number of participants, not the total number of measurements.  The R statistical program was used to conduct all analyses.[@R2018] All data and R code used in the meta-analyses is available [here](https://github.com/awconway/zhf-review).

@@

```{r fig-1,  fig.cap="PRISMA Flow Diagram", echo=FALSE, fig.width=5.5, fig.path="figures/", include=params$distill, eval=params$distill}
# https://github.com/awconway/ggprisma
ggprisma::ggprisma(retrieved = 131, included = 18, duplicates = 35, full_text = 23, wrong_intervention = 3, wrong_comparator = 2, wrong_design = 1, awaiting_classification = 2)
```

Detailed characteristics of included studies can be viewed on the [webpage](https://aconway.dev/zhf/) containing supplementary information for the systematic review. The primary comparison of zero-heat-flux versus core temperature measurements (eg, arterial, bladder, esophageal or rectal) consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. In total, data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included in this comparison (two studies did not report the total number of measurements included in their analysis). The sensitivity analysis for the primary comparison with only studies that were judged as low risk of bias across all domains included `r low_risk$Comparisons` comparisons from `r low_risk$Studies` studies that enrolled `r low_risk$Participants` participants with `r low_risk$Measurements` paired measurements (two studies did not report the total number of measurements included in their analysis). There were `r ICU$Studies` studies that compared zero-heat-flux to core temperature measurements in ICU patients, comprising `r ICU$Comparisons` comparisons with `r ICU$Measurements` measurements from `r ICU$Participants` participants. Another `r OT$Studies` studies were included that compared zero-heat-flux to core temperature measurements in patients undergoing surgery, comprising `r OT$Comparisons` comparisons with `r OT$Measurements` measurements from `r OT$Participants` participants. Nasopharyngeal thermometers were used as the comparator devices in `r NPA$Studies` studies, with `r NPA$Measurements` paired measurements from `r NPA$Measurements` participants. Sublingual thermometers were used as the comparator devices in `r SL$Studies` studies that reported `r SL$Measurements` paired measurements from `r SL$Participants` participants.

Most studies included in this systematic review evaluated the zero-heat-flux temperature monitoring system manufactured by 3M. Previously known as the SpotOn Temperature Monitoring System, the 3M zero-heat-flux device is now referred to commercially as the Bair Hugger Temperature Monitoring System. One study evaluated the Tcore^TM^ device.[@sastre2019evaluation] The oldest study included in this review, published in 2000, was an evaluation of the Coretemp device from Terumo.[@Harioka_2000] All comparisons reported adherence to the zero-heat-flux device manufacturer instructions and placed the sensor on the forehead of the participants. One included study also reported results for comparisons where the zero-heat-flux thermometer was not placed on the forehead. We did not include these comparisons in our meta-analysis.

@@

Table 1 presents results of the primary, sensitivity and all subgroup analyses for the primary comparison as well as secondary comparisons. The pooled estimate for the mean bias between zero-heat-flux and core temperature measurements was `r primary$bias_mean`°C. The population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` studies). The mean bias was again `r low_risk$bias_mean`°C with population limits of agreement spanning from `r low_risk$CI_L_rve`°C to `r low_risk$CI_U_rve`°C. A further sensitivity analysis excluding studies that received industry funding revealed population limits of agreement that were considerably wider than the primary analysis (`r conflict$CI_L_rve`°C to `r conflict$CI_U_rve`°C). The evidence rating for this sensitivity analysis was downgraded to low quality though, due to concerns about study limitations, inconsistency and imprecision.

We conducted two subgroup analyses for the primary comparison according to the clinical setting in which the study was conducted. In the subset of studies conducted in the ICU, the mean bias was `r ICU$bias_mean`°C with population limits of agreement between `r ICU$CI_L_rve`°C and `r ICU$CI_U_rve` °C. The GRADE rating for quality of evidence was downgraded to very low quality due to study limitations, inconsistency and imprecision. In the subset of studies that evaluated the use of zero-heat-flux thermometers during surgery found a mean bias of `r OT$bias_mean`°C with population limits of agreement from `r OT$CI_L_rve`°C to `r OT$CI_U_rve`°C. The quality of evidence was rated as moderate quality due to concerns about study limitations.

@@

Zero-heat-flux thermometers were compared with sublingual thermometers and nasopharyngeal thermometers in the studies included in this review. The mean bias between zero-heat-flux and nasopharyngeal temperature measurements in meta-analysis of results from `r NPA$Studies` studies was `r NPA$bias_mean`°C. Population limits of agreement were `r NPA$CI_L_rve`°C to `r NPA$CI_U_rve`°C. We downgraded the quality of evidence to low, again due to concerns about imprecision and study limitations. The mean bias between zero-heat-flux and sublingual temperature measurements in meta-analysis of results from `r SL$Studies` studies was `r SL$bias_mean`°C. Due to the limited number of studies and measurements, population limits of agreement were extremely wide, spanning outside physiologically plausible range for temperature. As such, they are not reported.  The quality of evidence for this comparison was rated as very low quality due to serious concerns about imprecision.

@@

It should be noted that we did not include studies that used the Temple Touch Pro because it is not strictly a zero-heat-flux device. This is a new thermometer that is similar to the zero-heat-flux device in that it is placed cutaneously, but the underlying technology is different.[@evron2017evaluation]

@@
